Recommendation Generation for Performance Improvement by using Cross-Organizational Process Mining

### 2015-09-07 4_methodology/methodology.tex
no message

>>>>

\subsection{Mismatch Pattern Analysis}
\label{subsec:mismatch-pattern-analysis}
In order to learn from other organizations, it is necessary to spot the differences between process models of different organizations. In this phase of methodology, differences between process models will be revealed by the mismatch patterns which are defined by Dijkman \cite{dijkman2007mismatch}. In the process mining stage, process models are mined to create Workflow nets; however, mismatch pattern definitions are closer to business environment and it is more suitable spot patterns in BPMN notation. Thus, mined process models are converted to BPMN models and then mismatch pattern analysis is applied. After spotting differences, mismatch patterns for activities and control flows are revealed for the organizations.

Business Process Modeling Notation (BPMN) is one of the widely accepted modeling language in the industry and its a standardized notation by the Object Management Group (OMG) since 2004. In this stage, BPMN formulation is used since mismatch patterns have roots in real-life business environment \cite{dijkman2007mismatch}.

For each organization, mined process models are converted to BPMN models and mismatch patterns analysis is undertaken on them, in order to locate differences between two process models or cluster of process models. In addition, since performance indicators are calculated based on a starting and ending points in the process model, same approach is applied to locate mismatch patterns. In other words, differences of process models are located through a starting activity to an ending activity. When the complete set of mismatches between process models are required, this approach is used with starting and ending points as source and sink activities. For each organization, mismatch patterns are gathered and stored for the further analysis as visualized in Figure~\ref{fig:mismatch-pattern-analysis-blackbox}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{4_methodology/mismatch-pattern-analysis-blackbox}
  \caption{Mismatch Pattern Analysis Stage as Black-box}
  \label{fig:mismatch-pattern-analysis-blackbox}
\end{figure}

\subsection{Recommendation Generation}
\label{subsec:recommendation-generation}
Recommendation generation stage in the methodology is the final and core stage where all information retrieved from event logs until now are utilized. In this study, idea of recommendation is based on providing a set of mismatch patterns for each organization so that they can enhance their processes. These mismatch patterns are generated by comparing the process models of other organizations, particularly which are performing better in terms of their performance indicator values. In addition, clustering of organizations is undertaken to generalize the way of identifying which organizations perform better in this environment.

Algorithm of recommendation generation function is based on the idea of checking other clusters for a significant change in performance indicators, where significance is defined by the threshold provided by user. After finding significant difference, all organizations in other clusters are checked against mismatch patterns with the starting and ending activities defined in performance indicators. With this constraining, only mismatches which are located between the activities that causes high level of difference in performance indicators are analyzed.

This stage of methodology can be visualized as gathering inputs of mismatch patterns data and cluster analysis data in Figure~\ref{fig:recommendation-generation-blackox}. In addition, a performance difference threshold is gathered from the user to specify how much difference between clusters is necessary to check for mismatch patterns and generate recommendations. This stage generates recommendation data for each organization which contains a set of mismatch patterns.
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{4_methodology/recommendation-generation-blackbox}
  \caption{Recommendation Generation Stage as Black-box}
  \label{fig:recommendation-generation-blackox}
\end{figure}

\subsection{Implementation in ProM Framework}
\label{subsec:implementation}
Methodology of this thesis study is implemented in ProM \cite{verbeek2010prom}, which is an extensible framework that supports a wide variety of process mining techniques in form of plugins. ProM is an open source framework for process mining algorithms which provides a platform to users and developers. Aim of this framework is to create a standard process mining platform that is accepted in industry and academia with an active community. Approach of this thesis study is implemented with its each stage as a standalone plugin that enables extensions for further studies. Developed set of plugins are packaged with the name of \textit{CrossOrgProcMin}\footnote{\url{http://www.promtools.org/prom6/packages/CrossOrgProcMin}} and published open-source\footnote{\url{http://github.com/onuryilmaz/cross-org-proc-min}} being available in the latest version of ProM release.

Approach of this study is divided into stand-alone stages where inputs and outputs between them are defined strictly. With the help of this understanding, each stage is developed as a stand-alone plug-in in ProM framework which can be called separately. For this aim, five plug-ins are developed which are visualized in Figure~\ref{fig:architecture} to provide a high-level perspective. In addition, a set of utilities are developed to visualize and persistence of data in ProM environment. 
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{4_methodology/architecture}
  \caption{Software Architecture for  Cross-Organizational Process Miner}
  \label{fig:architecture}
\end{figure}
